### 2022-03-08 papers/2022_Enschede_Talk/pres.i.tex
new stuff

\documentclass[fleqn]{beamer}

\mode<presentation>
{
  \usetheme{default}
%   \useinnertheme[shadow=true]{rounded}

  \useinnertheme{circles}

%    \useoutertheme{infolines}
  % or ...
  \setbeamersize{text margin left=1em,text margin right=1em}

%   \setbeamercovered{transparent}
  % or whatever (possibly just delete it)

  \beamertemplatenavigationsymbolsempty

% Display frame numbers in footline
  \setbeamertemplate{footline}[frame number]
}

\usepackage{etex}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[english]{babel}
\usepackage{amsmath}
% \usepackage{amsthm}
% \usepackage{stmaryrd}
\usepackage{lmodern}
\usepackage{mathpartir}

\usepackage{centernot}

\usepackage{colortbl}
\usepackage{multirow}

\usepackage[purexy]{qsymbols}
\usepackage{graphicx}

\usepackage{cprotect}
\usepackage{listings}
\usepackage{lstautogobble}
\usepackage{relsize}

\usepackage[noend]{algpseudocode}

\usepackage{ifthen}
\usepackage{pgf}
\usepackage{tikz}
\usetikzlibrary{scopes}
\usetikzlibrary{decorations}
  \usetikzlibrary{decorations.pathmorphing}
\usetikzlibrary{arrows}
\usetikzlibrary{automata}
\usetikzlibrary{positioning}
\usetikzlibrary{chains}
\usetikzlibrary{shapes.geometric}
\usetikzlibrary{shapes.callouts}
\usetikzlibrary{shapes.misc}
\usetikzlibrary{shapes.multipart}
\usetikzlibrary{fit}
\usetikzlibrary{calc}

\usepackage{pgfplots}

% \tikzstyle{stack}=[inner sep=0pt,minimum size=2mm]
% \tikzstyle{ssline}=[->,snake=snake,segment amplitude=.2mm,segment length=3mm,line after snake=1mm]
% \tikzstyle{fgnode}=[circle,draw,inner sep=0pt,minimum size=2mm]

\tikzset{
  -|-/.style={
    to path={
      (\tikztostart) -| ($(\tikztostart)!#1!(\tikztotarget)$) |- (\tikztotarget)
      \tikztonodes
    }
  },
  -|-/.default=0.5,
  |-|/.style={
    to path={
      (\tikztostart) |- ($(\tikztostart)!#1!(\tikztotarget)$) -| (\tikztotarget)
      \tikztonodes
    }
  },
  |-|/.default=0.5,
}

\tikzset{
  invisible/.style={opacity=0},
  visible on/.style={alt=#1{}{invisible}},
  alt/.code args={<#1>#2#3}{%
    \alt<#1>{\pgfkeysalso{#2}}{\pgfkeysalso{#3}} % \pgfkeysalso doesn't change the path
  },
}

% The goal is to translate
%    \overlaynode<red,blue>{hallo};
% into
%    \node[red]{hallo};
%    \node[blue]{hallo};
\makeatletter
\def\overlaynode<#1>#2;{
        \gdef\stacknodecommonpart{#2}
        \pgfkeys{/typeset node/.list={#1}}
        % we are lazy
        % pgfkeys will translate
        %    \pgfkeys{/typeset node/.list={red,blue}}
        % into
        %    \pgfkeys{/typeset node=red}
        %    \pgfkeys{/typeset node=blue}
}
\pgfkeys{
    /typeset node/.code={
        \edef\pgf@marshal{\noexpand\node[#1]\stacknodecommonpart;}
        \pgf@marshal
    }
}
% \tikz{
%     \overlaynode<red,{blue,xshift=1}>{{Hello}}; % notice the nested {{}}
% %   \overlaynode<red,{blue,xshift=1}>[]{Hello}; % workaround
%     \overlaynode<red,{blue,rotate=5}>at(1,0)[draw]{from};
%     \overlaynode<red,{blue,scale=1.1}>[circle]at(2,0)[draw]{the};
%     \overlaynode<red,{blue,opacity=.5}>[fill=yellow!50]at(3,0){other};
%     \overlaynode<{red,rectangle callout,fill},{blue,ellipse callout,draw}>at(4,0){side};
% }
% \tikz{
%     \overlaynode<
%             {fill=red,callout absolute pointer=(45:2)},
%             {fill=yellow,callout absolute pointer=(135:2),callout pointer shorten=1cm},
%             {fill=green,callout relative pointer=(-135:2),callout pointer width=.5cm},
%             {fill=blue,callout relative pointer=(-45:2),callout pointer shorten=1cm,callout pointer width=.5cm}
%         >
%         [rectangle callout,text=white]
%         at(0,0){can you}
%     ;
% }

\def\overlaynodedrawfill{\pgfutil@ifnextchar[{\overlaynodedrawfill@opt}{\overlaynodedrawfill@opt[]}}
\def\overlaynodedrawfill@opt[#1]<#2>#3;{
    \begin{scope}[transparency group,draw=black,fill=white,line cap=round,line join=round,#1]
        \pgfmathsetmacro\pgflinewidthdouble{2\pgflinewidth}
        \overlaynode<#2>[draw=pgfstrokecolor,line width=\pgflinewidthdouble]#3;
        \overlaynode<#2>[fill=pgffillcolor]#3;
    \end{scope}
}

\makeatother

% \tikz{
%     \overlaynodedrawfill[draw=magenta,fill=cyan,opacity=.5]<
%             {callout absolute pointer=(45:2)},
%             {callout absolute pointer=(135:2),callout pointer shorten=1cm},
%             {callout relative pointer=(-135:2),callout pointer width=.5cm},
%             {callout relative pointer=(-45:2),callout pointer shorten=1cm,callout pointer width=.5cm}
%         >
%         [rectangle callout]
%         at(0,0){hear me}
%     ;
% }
%
% \tikz{
%     \overlaynodedrawfill<
%             {cloud callout,callout absolute pointer=(90:2),inner sep=-20},
%             {rectangle callout,callout absolute pointer=(210:2),minimum height=30},
%             {ellipse callout,callout absolute  pointer=(-30:2)}
%         >
%         at(0,0){nice to meet you}
%     ;
% }
%
% \tikz{
%     \overlaynodedrawfill<
%             {starburst},
%             {cloud,inner sep=-10}
%         >
%         at(0,0){where you been}
%     ;
% }
%
% \usetikzlibrary{shapes.arrows}
% \tikz{
%     \overlaynodedrawfill[arrow box,text opacity=0,minimum height=50,minimum width=50,inner xsep=-10]<
%             {rotate=22.5},
%             {rotate=45},
%             {rotate=67.5},
%             {text opacity=1}
%         >
%         at(0,0){I can}
%     ;
% }
%

\usepackage{packages/isabelle}
\usepackage{packages/isabelletags}
\usepackage{packages/isabellesym}
\usepackage{packages/comment}

% \isabellestyle{it}

\def\isachardoublequote{}%
\def\isachardoublequoteopen{}%
\def\isachardoublequoteclose{}%

\newcommand{\isainnerkeyword}[1]{{\textbf{#1}}}
\newcommand{\isasymexistsA}{\isamath{\exists_{\textsc A}\,}}

\def\isadelimproof{}
\def\endisadelimproof{}
\def\isatagproof{}
\def\endisatagproof{}
\def\isafoldproof{}
\def\isadelimproof{}
\def\endisadelimproof{}

\def\isastylescript{\sl}%

\input{lstisabelle}
\newcommand{\isai}{\lstinline[language=isabelle,basicstyle=\normalsize\ttfamily\slshape]}

\newcommand\CC{C\nolinebreak[4]\hspace{-.05em}\raisebox{.3ex}{\relsize{-1}{\textbf{++}}}}

\newcommand{\eqdef}{\mathrel{{=}_{def}}}
\newcommand{\iffdef}{\mathrel{{\mathord{\iff}\!\!}_{def}}}

\makeatletter
\newcommand*{\overlaynumber}{\number\beamer@slideinframe}
\makeatother

\AtBeginSection[] % Do nothing for \section*
{
  \begin{frame}<beamer>
    \frametitle{Outline}
    \tableofcontents[currentsection]
  \end{frame}
}

\title{Refinement of Parallel Algorithms down to LLVM}

% \subtitle
% {Subtitle} % (optional)

\author[Peter Lammich]{Peter Lammich}
% - Use the \inst{?} command only if the authors have different
%   affiliation.

\institute[UoM] % (optional, but mostly needed)
{ University of Twente}
% - Use the \inst command only if there are several affiliations.
% - Keep it simple, no one is interested in your street address.

\date {Feb 2022}
% {2008-12-01}

% If you have a file called "university-logo-filename.xxx", where xxx
% is a graphic format that can be processed by latex or pdflatex,
% resp., then you can add a logo as follows:

% \pgfdeclareimage[height=0.5cm]{university-logo}{university-logo-filename}
% \logo{\pgfuseimage{university-logo}}

% Delete this, if you do not want the table of contents to pop up at
% the beginning of each subsection:

% If you wish to uncover everything in a step-wise fashion, uncomment
% the following command:

%\beamerdefaultoverlayspecification{<+->}

%\mathchardef\-="2D
%\renewcommand\-{\text{-}}

\newcommand{\mc}{\color{blue}}
\newcommand{\term}[1]{{\mc#1}}

\let\olddisplaystyle\displaystyle
\newcommand{\mydisplaystyle}{\olddisplaystyle\mc}
\let\displaystyle\mydisplaystyle

\newcommand{\smc}{\everymath{\mc}}
\smc

\lstset{basicstyle=\color{blue}}

% \newcommand<>{\btikzset}[2]{\alt#3{\tikzset{#1}}{\tikzset{#2}}}

\tikzset{onslide/.code args={<#1>#2}{%
  \only<#1>{\pgfkeysalso{#2}} % \pgfkeysalso doesn't change the path
}}

\tikzset{uncover/.code args={<#1>#2}{%
  \uncover<#1>{\pgfkeysalso{#2}} % \pgfkeysalso doesn't change the path
}}

\tikzset{>=latex}

\lstset{autogobble}

\newcommand{\natN}{{\text{nat}_{\mathord{<}N}}}

\newcommand{\high}[1]{{\color{blue}#1}}

\begin{document}
% \input{macros}

\begin{frame}
  \titlepage
\end{frame}

\newcommand{\insertsectitle}{}

\setbeamertemplate{frametitle}{\vspace{.7em}\insertframetitle\hfill \small\raisebox{10pt}{\insertsectitle}}

# The Isabelle Refinement Framework
  \definecolor{colA}{RGB}{11, 97,164}
  \definecolor{colB}{RGB}{204,246,  0}
  \definecolor{colC}{RGB}{166,  0,166}
  \definecolor{colD}{RGB}{255,146,  0}
  \definecolor{colTCB}{RGB}{255, 0, 0}

  \center
  \begin{tikzpicture}[node distance=.01\textwidth,outer sep=0pt,
  basicLayer/.style={
    draw=black,
    rounded corners,
    text width=.9\textwidth,
    font={\sffamily\bfseries\color{black}},
    align=center,
    minimum height=.03\textheight,
%    text height=12pt,
%    text depth=9pt
    },
  layerA/.style={basicLayer,fill=colA!30},
  layerB/.style={basicLayer,fill=colB!30},
  layerC/.style={basicLayer,fill=colC!30},
  layerD/.style={basicLayer,fill=colD!30},
  layerDD/.style={basicLayer,fill=colD!40,text width=.2\textwidth},
  layerTCB/.style={basicLayer,fill=colTCB!30,text width=.2\textwidth},
  layerTCBB/.style={basicLayer,fill=colTCB!40,text width=.2\textwidth},
  ]

  \node[layerA] (AlgDS) {Algorithms and Data Structures};

  \node[layerB, below=of AlgDS.south west, anchor=north west, text width=.437\textwidth] (ne) {NE-Monad};
  \node[layerA, right=of ne.north east, anchor=north west, text width=.433\textwidth, minimum height=.134\textheight] (llDS) {Low-Level Algorithms \\and Data Structures};

  \node[layerB, below=of ne.south west, anchor=north west, text width=.437\textwidth] (sepref) {Sepref};

  \node[layerC, below=of sepref.south west, anchor=north west] (sepl) {Program Logic and VCG};
  \node[layerD, below=of sepl.south west, anchor=north west, minimum height=.34\textheight] (sembox) {};

  \node[layerTCB, below left=of sembox.north east, minimum height=.32\textheight, text width=.537\textwidth] (tcbbox) {};

  \node[above right=of sembox.south west,font={\sffamily\bfseries\color{black}}] {Back End};

  \node[above left=of tcbbox.south east,font={\sffamily\bfseries\color{black}}] {Trusted Code Base};

%   \node[layerD, right=of sembox.west] (preproc) {Preprocessor};
%   \node[layerD, right=1cm of preproc] (instr) {Instructions};
%   \node[layerD, right=1cm of instr] (codegen) {Code Generator};

  \node[layerTCBB, below left=of tcbbox.north east] (codegen) {Code Generator};
  \node[layerTCBB, left=1cm of codegen] (instr) {Instructions};
  \node[layerDD, left=1cm of instr] (preproc) {Preprocessor};

  \node[layerTCBB, below=of instr] (mmodel) {Memory Model};
  \node[layerTCBB, below=of mmodel] (mmonad) {SE-Monad};

  \draw[->] (preproc) -- (instr);
  \draw[->] (instr) -- (codegen);

  \end{tikzpicture}

{

\usebackgroundtemplate{%
\only<3->{
\tikz\node[opacity=0.1] {\includegraphics[height=\paperheight]{isabelle-llvm.png}};
}
\only<2>{
\tikz\node[opacity=1] {\includegraphics[height=\paperheight]{isabelle-llvm.png}};
}
}

#! Isabelle LLVM Back End
  *<2-> Shallow embedding of small fragment of LLVM
    * just enough to express our programs
    * code generator translates to actual LLVM text
  *<3-> Simple memory model {\small
    \begin{lstlisting}
      datatype addr == ADDR (bidx: nat) (idx: nat)
      datatype ptr == PTR_NULL   |   PTR_ADDR (the_addr: addr)
      datatype val == LL_INT lint   |   LL_STRUCT val list   |   LL_PTR ptr

      datatype block == FRESH   |   FREED   |   is_alloc: ALLOC (vals: "val list")
      typedef memory == { \mu :: nat => block. finite {b. \mu b ~= FRESH} }
    \end{lstlisting}}

  *<4-> Using state-error monad
    \begin{lstlisting}
      \alpha llM = memory => (FAIL | SUCC (\alpha \x memory))
    \end{lstlisting}

}

#! Shallow Embedding Example
  \newcommand{\mknd}[1]{\makebox[0pt]{\tikz[remember picture]{\node (#1) {};}}}
  {\small\centering
  \begin{minipage}[t]{.35\textwidth}
    \begin{lstlisting}[escapechar=!]
      fib:: 64 word !\mknd{ndparams}!=> 64 word ll!\mknd{ndrettype}!M
      fib n = do {
        t <- ll_icmp_ule n 1;
        llc_if!\mknd{stmtif}! t
          (return n)
          (do {
            n_1 <- ll_sub!\mknd{stmtsub}! n 1;
            a   <- fib!\mknd{stmtrec}! n_1;
            n_2 <- ll_sub n!\mknd{ndargs}! 2;
            b   <-!\mknd{ndbind}! fib n_2;
            c   <- ll_add a b;
            return!\mknd{ndreturn}! c
          }) }
    \end{lstlisting}
%     \begin{tikzpicture}[remember picture,overlay,ND/.style={rectangle callout, anchor=west,draw , fill=red!50, callout absolute pointer={#1}}]
%       \node[ND=(ndrettype.east),anchor=south] at ($(ndrettype.north) + (0,.5cm)$) {\footnotesize state/error monad};
%       \pause
%       \node[ND=(ndparams.east)] at ($(ndparams -| 6,0) + (0,.5cm)$) {\footnotesize parameter types: words, pointers, pairs};
%       \pause
%       \node[ND=(stmtif.east)] at ($(stmtif -| 6,0) + (0,.5cm)$) {\footnotesize control flow (if, [optional: while])};
%       \pause
%       \node[ND=(stmtsub.east)] at ($(stmtsub -| 6,0)+(0,.5cm)$) {\footnotesize standard instructions};
%       \pause
%       \node[ND=(stmtrec.east)] at ($(stmtrec -| 6,0)+(0,.5cm)$) {\footnotesize function calls};
%       \pause
%       \node[ND=(ndargs.east)] at ($(ndargs -| 6,0)+(0,.5cm)$) {\footnotesize arguments: variables and constants};
%       \pause
%       \node[ND=(ndbind.east)] at ($(ndbind -| 6,0)+(0,.5cm)$) {\footnotesize monad: bind, return};
%     \end{tikzpicture}
    \begin{tikzpicture}[remember picture,overlay,ND/.style={anchor=west}]
      \pause
      \node[ND,anchor=south] at ($(ndrettype.north) + (0,.5cm)$) (X) {\footnotesize state/error monad};
      \draw (X) -- (ndrettype);

      \pause
      \node[ND] at ($(ndparams -| 6,0) + (0,.5cm)$) (X) {\footnotesize types: word, pointer, struct};
      \draw (X.west) -- (ndparams);

      \pause
      \node[ND=(ndbind.east)] at ($(ndbind -| 6,0)+(0,.5cm)$) (X) {\footnotesize monad: bind, return};
      \draw (X.west) -- (ndbind);
      \draw (X.west) -- (ndreturn);

      \pause
      \node[ND=(stmtsub.east)] at ($(stmtsub -| 6,0)+(0,.5cm)$) (X) {\footnotesize standard instructions (\is{ll_<opcode>})};
      \draw (X.west) -- (stmtsub);

      \pause
      \node[ND=(ndargs.east)] at ($(ndargs -| 6,0)+(0,.5cm)$) (X) {\footnotesize arguments: variables and constants};
      \draw (X.west) -- (ndargs);

      \pause
      \node[ND=(stmtif.east)] at ($(stmtif -| 6,0) + (0,.5cm)$) (X) {\footnotesize control flow (if, [optional: while])};
      \draw (X.west) -- (stmtif);

      \pause
      \node[ND=(stmtrec.east)] at ($(stmtrec -| 6,0)+(0,.5cm)$) (X) {\footnotesize function calls (rec.\ via fixp in ccpo)};
      \draw (X.west) -- (stmtrec);

    \end{tikzpicture}
  \end{minipage}
  }

#! Code Generation
  \uncover<2->{compiling control flow + pretty printing}

  {\small\centering
  \begin{minipage}[t]{.35\textwidth}
    \begin{lstlisting}
      fib:: 64 word => 64 word llM
      fib n = do {
        t <- ll_icmp_ule n 1;
        llc_if t

          (return n)
          (do {
            n_1 <- ll_sub n 1;
            a   <- fib n_1;
            n_2 <- ll_sub n 2;
            b   <- fib n_2;
            c   <- ll_add a b;
            return c
          }) }
    \end{lstlisting}
  \end{minipage}
  % \hspace*{.04\textwidth}
  \hfill\pause
  \begin{minipage}[t]{.57\textwidth}
    \begin{lstlisting}[language=LLVM, literate={}]
    define i64 @fib(i64 %n) {
      start:
        %t = icmp ule i64 %n, 1
        br i1 %t, label %then, label %else
      then:
        br label %ctd_if
      else:
        %n_1 = sub i64 %n, 1
        %a    = call i64 @fib (i64 %n_1)
        %n_2 = sub i64 %n, 2
        %b    = call i64 @fib (i64 %n_2)
        %c    = add i64 %a, %b
        br label %ctd_if
      ctd_if:
        %x1a = phi i64 [%n,%then], [%c,%else]
        ret i64 %x1a }
    \end{lstlisting}
  \end{minipage}
  }

#! Preprocessor
  * Only restricted terms accepted by code generator
    * good to keep code generation simple
    * tedious to write manually
  * Preprocessor transforms terms into restricted format
    * proves equality (via Isabelle kernel)
  * Motto: Keep TCB small, preprocessor makes it usable

#! Example: Preprocessing Euclid's Algorithm
  \small
  \begin{lstlisting}
  euclid :: 64 word => 64 word => 64 word
  euclid a b = do {
    (a,b) <- llc_while
      (%(a,b) => ll_cmp (a ~= b))
      (%(a,b) => if (a<=b) then return (a,b-a) else return (a-b,b))
      (a,b);
    return a }
  \end{lstlisting}
  \pause
  preprocessor defines function \is{euclid_0} and proves
  \begin{lstlisting}
  euclid a b = do {
      ab <- ll_insert_1 init a; ab <- ll_insert_2 ab b;
      ab <- euclid_0 ab;
      ll_extract_1 ab  }
  euclid_0 s = do {
    a <- ll_extract_1 s;
    b <- ll_extract_2 s;
    ctd <- ll_icmp_ne a b;
    llc_if ctd do {...; euclid_0 ...} }
  \end{lstlisting}

#! Reasoning about LLVM Programs
  *<+-> Separation Logic
    \begin{lstlisting}
      \<alpha> :: memory -> amemory :: sep_algebra
      wp c Q s == \<exists>r s'. c s = SUCC r s' \and Q r (\<alpha> s')
      {P} c {Q} == \<forall>F s. (P * F) (\<alpha> s) --> wp c (\<lambda>r s'. (Q r * F) s') s
    \end{lstlisting}

    * defined wrt.\ shallowly embedded semantics
    * proof rules are proved theorems!

  *<+-> Automation: VCG, frame inference, heuristics to discharge VCs
    * these prove theorems!

  *<+-> Basic Data Structures: signed/unsigned integers, Booleans, arrays

# Sepref
  * Semi-automatic translation of functional to imperative program
  * Data refinement to imperative DS
    * e.g. list to array
  * Proves refinement theorem

#! Example: Binary Search
  \newcommand{\mknd}[1]{\makebox[0pt]{\tikz[remember picture]{\node (#1) {};}}}
  \begin{lstlisting}
    definition "bin_search xs x = do {
      (l,h) <- while (bin_search$\mknd{invar}$_invar xs x)
        (%(l,h). l<h)
        (%(l,h). do {
          ass$\mknd{assn}$ert (l<|xs| \<and> h<=|xs| \<and> l<=h);
          let m = l +$\mknd{midc}$ (h-l) div 2;
          if xs!m < x then return (m+1,h) else return (l,m)
        })
        (0,|xs|);
      return l
    }"
  \end{lstlisting}

  \onslide<2->

  \begin{tikzpicture}[remember picture,overlay,ND/.style={anchor=west}]
    \only<2->{
    \node[ND,anchor=west] at ($(invar) + (.5cm,-.5cm)$) (X) {\footnotesize invariant annotation};
    \draw (X.west) -- (invar);
    }

    \only<3->{
    \node[ND,anchor=west] at ($(assn) + (3cm,-3cm)$) (X) {\footnotesize hint for subsequent refinement};
    \draw (X.west) -- (assn);
    }

    \only<4->{
    \node[ND,anchor=west] at ($(midc) + (2,-1cm)$) (X) {\footnotesize overflow-safe midpoint computation};
    \draw (X.west) -- (midc);
    }
  \end{tikzpicture}

  \onslide<5->

  \begin{lstlisting}
  lemma bin_search_correct:
    sorted xs ==> bin_search xs x <= spec i. i=find_index (<=y) xs
  \end{lstlisting}

#! Example: Binary Search --- Refinement
  \newcommand{\mknd}[1]{\makebox[0pt]{\tikz[remember picture]{\node (#1) {};}}}
  \onslide<1->

  \begin{lstlisting}[escapechar=!]
  sepref_def bin_search$_\dagger$ is bin_search
    :: "(array$_A$!\mknd{assn1}! int$_A^{64}$)$^k$ $$* (int!\mknd{assn2}!$_A^{64}$)$^k$ -> int!\mknd{assn3}!$_A^{64}$
    unfolding bin_search_def
    apply (rule hfref_with_rdomI,!\mknd{assn4}! annot_snat_const "64")
    by sep!\mknd{auto}!ref
  \end{lstlisting}
  \onslide<2->

  \begin{tikzpicture}[remember picture,overlay,ND/.style={anchor=west}]
  \only<2-3>{
    \only<2->{
    \node[ND,anchor=south] at ($(assn1) + (4cm,-3cm)$) (X) {\footnotesize hints for data refinement};
    \draw (X) -- (assn1);
    \draw (X) -- (assn2);
    \draw (X) -- (assn3);
    \draw (X) -- (assn4);
    }

    \only<3->{
    \node[ND] at ($(auto) + (0,-2cm)$) (X) {\footnotesize automatic synthesis + proof};
    \draw (X.west) -- (auto);
    }
  }
  \end{tikzpicture}

  \onslide<4->
  proves: \is{(bin_search$_\dagger$, bin_search) \<in> (array$_A$ int$_A^{64}$)$^k$ * int${_A^{64}}^k$ -> int$_A^{64}$}
  \onslide<5->

  Combination with \is{bin_search_correct} yields:
  \begin{lstlisting}
  theorem bin_search$_\dagger$_correct:
    {(array$_A$ int$_A^{64}$ xs xs$_\dagger$ * int$_A^{64}$ x x$_\dagger$ * sorted xs)}
      (bin_search$_\dagger$ xs$_\dagger$ x$_\dagger$)
    {%i$_\dagger$. \<exists>i. array$_A$ int$_A^{64}$ xs xs$_\dagger$ * int$_A^{64}$ x x$_\dagger$ * int$_A^{64}$ i i$_\dagger$ * i=find_index (<=y) xs}
  \end{lstlisting}

#! Example: Binary Search --- Generated Code
  \newcommand{\mknd}[1]{\makebox[0pt]{\tikz[remember picture]{\node (#1) {};}}}

  \begin{lstlisting}[escapechar=!]
  export_llvm bin_se!\mknd{isaconst}!arch$_\dagger$ is int64_t bin_!\mknd{csig}!search(larray_t, elem_t)
  defines
    typedef !\mknd{tdefs}!int64_t elem_t;
    typedef struct { int64_t len; elem_t *data; } larray_t;
  file "!\mknd{file}!code/bin_search.ll"
  \end{lstlisting}

  \begin{tikzpicture}[remember picture,overlay,ND/.style={anchor=west}]
  \only<2-5>{
    \only<2->{
    \node[ND,anchor=south] at ($(isaconst) + (0,-3cm)$) (X) {\footnotesize Isabelle constant};
    \draw (X) -- (isaconst);
    }

    \only<3->{
    \node[ND] at ($(csig) + (0,-3cm)$) (X) {\footnotesize LLVM name and signature};
    \draw (X.west) -- (csig);
    }

    \only<4->{
    \node[ND] at ($(tdefs) + (0,-3cm)$) (X) {\footnotesize C-style typedefs};
    \draw (X.west) -- (tdefs);
    }

    \only<5->{
    \node[ND] at ($(file) + (0,-3cm)$) (X) {\footnotesize target file};
    \draw (X.west) -- (file);
    }
  }
  \end{tikzpicture}

  \onslide<6->

  Produces LLVM code and header file:
  \begin{lstlisting}[language=C]
    typedef int64_t elem_t;
    typedef struct {
      int64_t len;
      elem_t*data;
    } larray_t;

    int64_t bin_search(larray_t,elem_t);
  \end{lstlisting}

{
\usebackgroundtemplate{%
\only<3->{
\tikz\node[opacity=0.1] {\includegraphics[height=\paperheight]{isabelle-llvm-par.png}};
}
\only<2>{
\tikz\node[opacity=1] {\includegraphics[height=\paperheight]{isabelle-llvm-par.png}};
}
}

#![t] Isabelle-LLVM Parallel
  \newcommand{\mknd}[1]{\makebox[0pt]{\tikz[remember picture]{\node (#1) {};}}}

  *<2-> Add parallel operator: $c_1 \parallel c_2$
    *<3-> execute in parallel, fail on data race
  *<4-> Shallow embedding: make program report memory accesses

    \only<4>{
    \begin{tikzpicture}[node distance = 1cm, auto]
      \node (m) {$\mu$};
      \node[below left = of m] (m1) {$\rho_1,\mu_1$};
      \node[below right = of m] (m2) {$\rho_2,\mu_2$};
      \node[below right = of m1] (m') {$\mu'$};

      \draw[->] (m) -- node[swap] {\small $c_1$}  (m1);
      \draw[->] (m) -- node {\small $c_2$}  (m2);

      \draw[->] (m1) -- (m');
      \draw[->] (m2) -- (m');

      \node[above = .7em of m'] {\tiny combine};
      \node[above = 2em of m'] {\tiny check race};

    \end{tikzpicture}
    }

    \begin{onlyenv}<5-6>
    \begin{lstlisting}
      (c_1 || c_2) \mu ==
        (r_1,\rho$\mknd{r1}$_1,\mu_1) <- c_1 \mu --"\rightcomment{execute first strand}"
        (r_2,\rho$\mknd{r2}$_2,\mu_2) <- c_2 \mu --"\rightcomment{execute second strand}"
        assert no_race \rho_1 \rho_2 --"\rightcomment{fail on data race}"
        \mu' = combine \rho_1 \mu_1   \rho_2 \mu_2 --"\rightcomment{combine states}"
        return ((r_1,r_2), \rho_1 \union \rho_2, \mu')

    \end{lstlisting}

    \begin{tikzpicture}[remember picture,overlay,ND/.style={anchor=west}]
      \only<6->{
      \node[ND,anchor=south] at ($(r1) + (3cm,-.1cm)$) (X) {\footnotesize Access report};
      \draw (X) -- (r1);
      \draw (X) -- (r2);
      }
    \end{tikzpicture}
    \end{onlyenv}

}

#! Memory Allocation
  *<+-> Currently: deterministic semantics
    * $c_1~\mu$ and $c_2~\mu$ will allocate same memory
    * cannot be combined!

  *<+-> Use nondeterminism
    \begin{lstlisting}
      \alpha llM = memory => FAIL | SUCC ((\alpha \x report \x memory) set)
    \end{lstlisting}
    * malloc nondeterministically allocates some free address
    * on combination: exclude infeasible possibilities

  \onslide<+->
  \begin{lstlisting}
    (c_1 || c_2) \mu ==
      (r_1,\rho_1,\mu_1) <- c_1 \mu
      (r_2,\rho_2,\mu_2) <- c_2 \mu
      assume \rho_1.alloc \<inter> \rho_2.alloc = \emptyset --"\rightcomment{ignore infeasible combinations}"
      assert no_race \rho_1 \rho_2 --"\rightcomment{fail on data race}"
      \mu' = combine \rho_1 \mu_1   \rho_2 \mu_2 --"\rightcomment{combine states}"
      return ((r_1,r_2), \rho_1 \union \rho_2, \mu')

  \end{lstlisting}

#! Invariants
  * We prove for \is{llM} (enforced by subtype)
    * access reports are consistent with observed changes in memory
    * there is at least one possible result (no magic happens)

  * Sanity check for semantics

  * Allows us to prove symmetry of $\parallel$
    \begin{lstlisting}
      c_1 || c_2  =  swapres (c_2 || c_1)

      swapres m == (r_1,r_2)<-m; return (r_2,r_1)
    \end{lstlisting}

#! Code generator
  *<+-> We add \is{llc_par f_1 f_2 x_1 x_2 == f_1 x_1 || f_2 x_2}
    * \is{f_1, f_2} must be functions
  *<+-> Code generator generates
    * type casting boilerplate
    * call to external \is{parallel} function

    {\small
    \begin{lstlisting}[language=C++]
    void parallel(void (*f1)(void*), void (*f2)(void*), void *x1, void *x2)
    \end{lstlisting}
    }
  *<+-> For example, implemented using TBB:
    {\small
    \begin{lstlisting}[language=C++]
    {
      tbb::parallel_invoke([=]{f1(x1);}, [=]{f2(x2);});
    }
    \end{lstlisting}
    }

#! Amending higher layers of IRF
  *<+-> Prove concurrency rule
    \begin{lstlisting}
      ${}$        {P_1} c_1 {Q_1}    \and    {P_2} c_2 {Q_2}
    \end{lstlisting}

    \vspace*{-.8em}
    \begin{lstlisting}
      ==> {P_1 * P_2} c_1 || c_2 {%(r_1,r_2). Q_1 r_1 * Q_2 r_2}
    \end{lstlisting}

  *<+-> Sepref refines sequential to parallel execution
    \begin{lstlisting}
      npar f_1 f_2 x_1 x_2 == r_1 <- f_1 x_1; r_2 <- f_2 x_2; return (r_1,r_2)
    \end{lstlisting}
    refined to \is{llc_par}.

  *<+-> Backwards compatible with sequential Sepref!
    * Easy porting of existing algorithms

#! Parallel Quicksort (basic)

  \begin{lstlisting}
    psort xs ==
      if |xs|<=1 then return xs    --"\rightcomment{trivially sorted}"
      else
        (xs,m) <- partition_spec xs;  --"\rightcomment{partition}"
        (_,xs) <- with_split m xs (%xs_1 xs_2.
          npar psort psort xs_1 xs_2 --"\rightcomment{recursively sort partitions}"
        );
        return xs

    with_split i xs f ==
      assert (i < |xs|);       --"\rightcomment{split point must be in list}"
      (xs_1,xs_2) <- f (take i xs) (drop i xs); --"\rightcomment{execute f with halfs}"
      assert (|xs_1| = i \and |xs_2| = |xs| - i); --"\rightcomment{length of halfs must not change}"
      return (xs_1@xs_2) --"\rightcomment{return both halfs}"

  \end{lstlisting}

#! Parallel Quicksort (refined)
  \small
  \begin{lstlisting}
    psort xs n ==
      assert n=|xs|;
      if n<=1 then return xs
      else psort_aux xs n (log2 n * 2) --"\rightcomment{recursion depth limit}"

    psort_aux xs n d ==
      assert n=|xs|  --"\rightcomment{extra parameter for length}"
      if d=0 \or n<100000 then sort_spec xs --"\rightcomment{fallback to seq-sort}"
      else
        (xs,m) <- partition_spec xs;
        let bad = m<n div 8 \or (n-m < n div 8) --"\rightcomment{check unbalanced partition}"
        (_,xs) <- with_split m xs (%xs_1 xs_2.
          if bad then --"\rightcomment{sequentially recurse for unbalanced}"
            nseq psort_aux psort_aux (xs_1,m,d-1) (xs_2,n-m,d-1)
          else --"\rightcomment{recurse in parallel for balanced}"
            npar psort_aux psort_aux (xs_1,m,d-1) (xs_2,n-m,d-1)
        );
        return xs"

  \end{lstlisting}

#![t] Parallel Quicksort (Sepref + Code Export)
  \small

  *<1-> Sepref generates imperative program
    * using existing sequential \is{pdqsort} for fallback
    * using (new) sampling partitioner (proved correct + refined separately)

  *<2-> Correctness theorem:
    {\only<3->{\tiny}
    \begin{lstlisting}
    {arr$_A$ xs xs$_\dagger$ * idx$_A$ n n$_\dagger$ * n = |xs|}
      (psort$_\dagger$ xs$_\dagger$ n$_\dagger$)
    {%r. r=xs$_\dagger$ * \<exists> xs'. arr$_A$ xs' xs$_\dagger$ * sorted xs' * mset xs' = mset xs}
    \end{lstlisting}}
  *<3-> Instantiation to concrete weak ordering + code export
    {\only<4->{\tiny}
    \begin{lstlisting}
      interpretation unat: pcmp (%_. <) (%_. ll_icmp_ult) unat$^{64}_A$ \<proof>
      interpretation str: pcmp (%_. <) (%_. strcmp) str$^{64}_A$ \<proof>

      export_llvm
        unat.psort$_\dagger$ is "uint64_t* psort(uint64_t*, int64_t)"
        str.psort$_\dagger$ is "llstring* str_psort(llstring*, int64_t)"
        defines
          typedef struct {int64_t sz; struct {int64_t cap; char *data;};} llstring;
        file psort.ll
    \end{lstlisting}}

  *<4-> Link against C++ benchmark driver
    \begin{lstlisting}
      clang++ [...] lib_isabelle_llvm.cpp psort.ll benchmark.cpp
    \end{lstlisting}

\tikzset{myaxis/.style={
      ymajorgrids,
      title style={at={(0.2,.85)}},
      ybar=0pt,
      width=\textwidth,
      height=.85\textheight,
      bar width=3pt,
      xtick=data,
      symbolic x coords={rev-sorted-end-10,rev-sorted-end-1,sorted-end-.1,almost-sorted-50,random-boolean,organ-pipe,sorted-end-10,equal,rev-sorted-middle-.1,rev-sorted,sorted-middle-1,rev-sorted-middle-10,random,almost-sorted-.1,sorted,rev-sorted-middle-1,sorted-middle-.1,almost-sorted-10,almost-sorted-1,sorted-middle-10,rev-sorted-end-.1,sorted-end-1,random-dup-10,},
      xlabel near ticks, x tick label style={rotate=45,anchor=east,font=\tiny},
      legend style = {
%         legend pos= north east,
      at = {(.87,.97)},
%       cells={anchor=west},
      font=\scriptsize
      }
}}

# Benchmarks

  \begin{tikzpicture}
    \begin{axis}[
      title={\large Laptop (6 core), uint64},
      ymajorgrids,
      title style={at={(0.2,.85)}},
      ybar=0pt,
      width=\textwidth,
      height=.85\textheight,
      bar width=3pt,
      xtick=data,
      symbolic x coords={rev-sorted-end-10,rev-sorted-end-1,sorted-end-.1,almost-sorted-50,random-boolean,organ-pipe,sorted-end-10,equal,rev-sorted-middle-.1,rev-sorted,sorted-middle-1,rev-sorted-middle-10,random,almost-sorted-.1,sorted,rev-sorted-middle-1,sorted-middle-.1,almost-sorted-10,almost-sorted-1,sorted-middle-10,rev-sorted-end-.1,sorted-end-1,random-dup-10,},
      xlabel near ticks, x tick label style={rotate=45,anchor=east,font=\tiny},
      legend style = {
%         legend pos= north east,
      at = {(.87,.97)},
%       cells={anchor=west},
      font=\scriptsize
      }
    ]
      \input{data_laptop_uint64}
    \end{axis}
  \end{tikzpicture}

# Benchmarks

  \begin{tikzpicture}
    \begin{axis}[
      title={\large Laptop (6 core), string},
      ymajorgrids,
      title style={at={(0.2,.85)}},
      ybar=0pt,
      width=\textwidth,
      height=.85\textheight,
      bar width=3pt,
      xtick=data,
      symbolic x coords={rev-sorted-end-10,rev-sorted-end-1,sorted-end-.1,almost-sorted-50,random-boolean,organ-pipe,sorted-end-10,equal,rev-sorted-middle-.1,rev-sorted,sorted-middle-1,rev-sorted-middle-10,random,almost-sorted-.1,sorted,rev-sorted-middle-1,sorted-middle-.1,almost-sorted-10,almost-sorted-1,sorted-middle-10,rev-sorted-end-.1,sorted-end-1,random-dup-10,},
      xlabel near ticks, x tick label style={rotate=45,anchor=east,font=\tiny},
      legend style = {
%         legend pos= north east,
      at = {(.87,.97)},
%       cells={anchor=west},
      font=\scriptsize
      }
    ]
      \input{data_laptop_llstring}
    \end{axis}
  \end{tikzpicture}

# Benchmarks

  \begin{tikzpicture}
    \begin{axis}[
      title={\large Server (22 core), uint64},
      ymajorgrids,
      title style={at={(0.2,.85)}},
      ybar=0pt,
      width=\textwidth,
      height=.85\textheight,
      bar width=3pt,
      xtick=data,
      symbolic x coords={rev-sorted-end-10,rev-sorted-end-1,sorted-end-.1,almost-sorted-50,random-boolean,organ-pipe,sorted-end-10,equal,rev-sorted-middle-.1,rev-sorted,sorted-middle-1,rev-sorted-middle-10,random,almost-sorted-.1,sorted,rev-sorted-middle-1,sorted-middle-.1,almost-sorted-10,almost-sorted-1,sorted-middle-10,rev-sorted-end-.1,sorted-end-1,random-dup-10,},
      xlabel near ticks, x tick label style={rotate=45,anchor=east,font=\tiny},
      legend style = {
%         legend pos= north east,
      at = {(.87,.97)},
%       cells={anchor=west},
      font=\scriptsize
      }
    ]
      \input{data_server_uint64}
    \end{axis}
  \end{tikzpicture}

# Benchmarks

  \begin{tikzpicture}
    \begin{axis}[
      title={\large Server (22 core), string},
      ymajorgrids,
      title style={at={(0.2,.85)}},
      ybar=0pt,
      width=\textwidth,
      height=.85\textheight,
      bar width=3pt,
      xtick=data,
      symbolic x coords={rev-sorted-end-10,rev-sorted-end-1,sorted-end-.1,almost-sorted-50,random-boolean,organ-pipe,sorted-end-10,equal,rev-sorted-middle-.1,rev-sorted,sorted-middle-1,rev-sorted-middle-10,random,almost-sorted-.1,sorted,rev-sorted-middle-1,sorted-middle-.1,almost-sorted-10,almost-sorted-1,sorted-middle-10,rev-sorted-end-.1,sorted-end-1,random-dup-10,},
      xlabel near ticks, x tick label style={rotate=45,anchor=east,font=\tiny},
      legend style = {
%         legend pos= north east,
      at = {(.87,.97)},
%       cells={anchor=west},
      font=\scriptsize
      }
    ]
      \input{data_server_llstring}
    \end{axis}
  \end{tikzpicture}

# Speedup
  \begin{tikzpicture}
    \begin{axis}[
      xlabel near ticks,
      legend style = {
        at = {(.87,.3)},
        cells={anchor=west},
        font=\scriptsize
      },
      ymajorgrids,
      title={\large Laptop},
      title style={at={(0.15,.85)}},
      width=\textwidth,
      height=\textheight
    ]
      \input{sudata_laptop}

      \draw[black, thin, sharp plot]
      (axis cs:\pgfkeysvalueof{/pgfplots/xmin},\pgfkeysvalueof{/pgfplots/xmin}) --
      (axis cs:\pgfkeysvalueof{/pgfplots/xmax},\pgfkeysvalueof{/pgfplots/xmax});

%       \addplot[black, thin, sharp plot, update limits=false] {\x};
    \end{axis}
  \end{tikzpicture}

# Speedup
  \begin{tikzpicture}
    \begin{axis}[
      xlabel near ticks,
      legend style = {
        at = {(.87,.3)},
        cells={anchor=west},
        font=\scriptsize
      },
      ymajorgrids,
      title={\large Server},
      title style={at={(0.15,.85)}},
      width=\textwidth,
      height=\textheight
    ]
      \input{sudata_server}

      \draw[black, thin, sharp plot]
      (axis cs:\pgfkeysvalueof{/pgfplots/xmin},\pgfkeysvalueof{/pgfplots/xmin}) --
      (axis cs:\pgfkeysvalueof{/pgfplots/xmax},\pgfkeysvalueof{/pgfplots/xmax});

%       \addplot[black, thin, sharp plot, update limits=false] {\x};
    \end{axis}
  \end{tikzpicture}

{
\usebackgroundtemplate{%
\tikz\node[opacity=0.1] {\includegraphics[height=\paperheight]{isabelle-llvm-par.png}};
}

# Conclusion
  * Verification of parallel programs
    * stepwise refinement to tackle complexity
    * down to LLVM, small TCB
    * \high{fast} verified programs

  * Idea: shallow embedding, using access reports
    * backwards compatible with sequential IRF

  * Future work
    * state-of-the-art parallel sorting
    * fractional separation logic
    * more concurrency
    * complexity of parallel algorithms
    * GP-GPUs

}

\end{document}