### 2022-05-12 papers/ITP2022/sudata_laptop.tex
renaming files

\addplot+[color=red, mark color=black, mark options={fill=red}] coordinates {
  (1, 1.00)
  (2, 1.80)
  (3, 2.45)
  (4, 2.96)
  (5, 3.25)
  (6, 3.48)
  (7, 3.61)
  (8, 3.69)
  (9, 3.77)
  (10, 3.92)
  (11, 4.00)
  (12, 4.10)
};
\addlegendentry{verified};

\addplot+[color=black, mark color=black, mark options={fill=black}] coordinates {
  (1, 1.00)
  (2, 1.91)
  (3, 2.69)
  (4, 3.43)
  (5, 4.03)
  (6, 4.29)
  (7, 4.42)
  (8, 4.53)
  (9, 4.62)
  (10, 4.72)
  (11, 4.74)
  (12, 4.80)
};
\addlegendentry{sample sort};

\addplot+[color=blue, mark color=black, mark options={fill=blue}] coordinates {
  (1, 1.00)
  (2, 1.78)
  (3, 2.42)
  (4, 2.98)
  (5, 3.26)
  (6, 3.48)
  (7, 3.59)
  (8, 3.75)
  (9, 3.93)
  (10, 3.97)
  (11, 4.16)
  (12, 4.23)
};
\addlegendentry{std::sort(par-unseq)};