### 2023-03-23 papers/JAR_SI_ITP2022/sadata_llstring_laptop.tex

\addplot+[color=red, mark color=black, mark options={fill=red}] coordinates {
  (1000, 0.07)
  (2783, 0.26)
  (7743, 0.85)
  (21544, 2.71)
  (59948, 8.49)
  (166810, 16.94)
  (464159, 33.00)
  (1291550, 96.47)
  (3593814, 295.55)
  (10000000, 898.21)
};
\addlegendentry{verified};

\addplot+[color=black, mark color=black, mark options={fill=black}] coordinates {
  (1000, 0.08)
  (2783, 0.26)
  (7743, 0.81)
  (21544, 2.51)
  (59948, 8.00)
  (166810, 6.38)
  (464159, 25.89)
  (1291550, 86.31)
  (3593814, 294.83)
  (10000000, 962.54)
};
\addlegendentry{sample sort};

\addplot+[color=blue, mark color=black, mark options={fill=blue}] coordinates {
  (1000, 0.06)
  (2783, 0.10)
  (7743, 0.20)
  (21544, 0.49)
  (59948, 1.23)
  (166810, 5.23)
  (464159, 23.66)
  (1291550, 82.55)
  (3593814, 300.77)
  (10000000, 1015.65)
};
\addlegendentry{std::sort};