### 2023-03-23 papers/JAR_SI_ITP2022/sadata_llstring_server.tex

\addplot+[color=red, mark color=black, mark options={fill=red}] coordinates {
  (1000, 0.18)
  (2783, 0.25)
  (7743, 0.84)
  (21544, 2.71)
  (59948, 8.61)
  (166810, 16.24)
  (464159, 27.09)
  (1291550, 71.66)
  (3593814, 163.82)
  (10000000, 416.31)
};
\addlegendentry{verified};

\addplot+[color=black, mark color=black, mark options={fill=black}] coordinates {
  (1000, 0.17)
  (2783, 0.56)
  (7743, 0.85)
  (21544, 2.79)
  (59948, 8.62)
  (166810, 31.70)
  (464159, 39.46)
  (1291550, 86.36)
  (3593814, 189.04)
  (10000000, 470.72)
};
\addlegendentry{sample sort};

\addplot+[color=blue, mark color=black, mark options={fill=blue}] coordinates {
  (1000, 0.13)
  (2783, 0.19)
  (7743, 0.38)
  (21544, 0.63)
  (59948, 0.90)
  (166810, 2.14)
  (464159, 7.00)
  (1291550, 34.22)
  (3593814, 128.24)
  (10000000, 461.74)
};
\addlegendentry{std::sort};