### 2023-03-23 papers/JAR_SI_ITP2022/sadata_uint64_laptop.tex

\addplot+[color=red, mark color=black, mark options={fill=red}] coordinates {
  (10000, 0.48)
  (27826, 1.46)
  (77426, 4.05)
  (215443, 4.71)
  (599484, 9.13)
  (1668101, 19.85)
  (4641589, 45.21)
  (12915497, 138.15)
  (35938137, 391.04)
  (100000000, 1206.60)
};
\addlegendentry{verified};

\addplot+[color=black, mark color=black, mark options={fill=black}] coordinates {
  (10000, 0.34)
  (27826, 0.96)
  (77426, 0.95)
  (215443, 1.81)
  (599484, 4.12)
  (1668101, 12.44)
  (4641589, 40.17)
  (12915497, 122.38)
  (35938137, 352.97)
  (100000000, 1040.94)
};
\addlegendentry{sample sort};

\addplot+[color=blue, mark color=black, mark options={fill=blue}] coordinates {
  (10000, 0.09)
  (27826, 0.23)
  (77426, 0.61)
  (215443, 1.78)
  (599484, 5.47)
  (1668101, 16.12)
  (4641589, 51.32)
  (12915497, 159.44)
  (35938137, 491.69)
  (100000000, 1426.85)
};
\addlegendentry{std::sort};