### 2023-03-23 papers/JAR_SI_ITP2022/sadata_uint64_server.tex

\addplot+[color=red, mark color=black, mark options={fill=red}] coordinates {
  (10000, 0.39)
  (27826, 1.21)
  (77426, 3.69)
  (215443, 11.12)
  (599484, 15.01)
  (1668101, 19.28)
  (4641589, 34.43)
  (12915497, 84.48)
  (35938137, 232.61)
  (100000000, 733.26)
};
\addlegendentry{verified};

\addplot+[color=black, mark color=black, mark options={fill=black}] coordinates {
  (10000, 0.58)
  (27826, 0.94)
  (77426, 13.99)
  (215443, 17.99)
  (599484, 21.07)
  (1668101, 29.71)
  (4641589, 69.88)
  (12915497, 176.65)
  (35938137, 377.31)
  (100000000, 765.52)
};
\addlegendentry{sample sort};

\addplot+[color=blue, mark color=black, mark options={fill=blue}] coordinates {
  (10000, 0.14)
  (27826, 0.31)
  (77426, 0.74)
  (215443, 1.08)
  (599484, 2.84)
  (1668101, 7.83)
  (4641589, 25.13)
  (12915497, 95.56)
  (35938137, 277.38)
  (100000000, 821.05)
};
\addlegendentry{std::sort};