### 2023-03-23 papers/JAR_SI_ITP2022/sudata_llstring_laptop.tex

\addplot+[color=red, mark color=black, mark options={fill=red}] coordinates {
  (1, 1.00)
  (2, 1.83)
  (3, 2.52)
  (4, 3.03)
  (5, 3.42)
  (6, 3.67)
  (7, 3.92)
  (8, 3.95)
  (9, 4.20)
  (10, 4.19)
  (11, 4.54)
  (12, 4.65)
};
\addlegendentry{verified};

\addplot+[color=black, mark color=black, mark options={fill=black}] coordinates {
  (1, 1.00)
  (2, 1.92)
  (3, 2.72)
  (4, 3.37)
  (5, 3.90)
  (6, 4.17)
  (7, 4.30)
  (8, 4.43)
  (9, 4.58)
  (10, 4.69)
  (11, 4.71)
  (12, 4.79)
};
\addlegendentry{sample sort};

\addplot+[color=blue, mark color=black, mark options={fill=blue}] coordinates {
  (1, 1.00)
  (2, 1.75)
  (3, 2.37)
  (4, 2.84)
  (5, 3.17)
  (6, 3.18)
  (7, 3.48)
  (8, 3.52)
  (9, 3.74)
  (10, 3.87)
  (11, 4.08)
  (12, 4.14)
};
\addlegendentry{std::sort};