### 2023-03-23 papers/JAR_SI_ITP2022/sudata_uint64_laptop.tex

\addplot+[color=red, mark color=black, mark options={fill=red}] coordinates {
  (1, 1.00)
  (2, 1.81)
  (3, 2.68)
  (4, 3.52)
  (5, 4.35)
  (6, 5.15)
  (7, 5.51)
  (8, 5.95)
  (9, 6.14)
  (10, 6.34)
  (11, 6.87)
  (12, 6.72)
};
\addlegendentry{verified};

\addplot+[color=black, mark color=black, mark options={fill=black}] coordinates {
  (1, 1.00)
  (2, 1.92)
  (3, 2.84)
  (4, 3.73)
  (5, 4.55)
  (6, 5.31)
  (7, 5.79)
  (8, 6.20)
  (9, 6.29)
  (10, 6.42)
  (11, 6.60)
  (12, 6.85)
};
\addlegendentry{sample sort};

\addplot+[color=blue, mark color=black, mark options={fill=blue}] coordinates {
  (1, 1.00)
  (2, 1.93)
  (3, 2.86)
  (4, 3.79)
  (5, 4.70)
  (6, 5.55)
  (7, 5.99)
  (8, 6.38)
  (9, 6.74)
  (10, 7.11)
  (11, 7.35)
  (12, 7.61)
};
\addlegendentry{std::sort};