### 2024-03-23 benchmarks/sorting/distr_report.tex
better reproducible benchmark scripts

\documentclass[a4]{article}

\usepackage[T1]{fontenc}
\usepackage{color}

\usepackage{pgfplots}
\usepgfplotslibrary{groupplots}

\begin{document}

  \newcommand{\plottim}[3]{
  \begin{tikzpicture}
    \begin{axis}[
      ymajorgrids,
      title={\large #1},
      title style={at={(0.12,.75)}},
      ybar=0pt,
      width=\textwidth,
      height=.249\textheight,
      bar width=2.5pt,
      xtick=data,
      symbolic x coords={rev-sorted-end-10,rev-sorted-end-1,sorted-end-.1,almost-sorted-50,random-boolean,organ-pipe,sorted-end-10,equal,rev-sorted-middle-.1,rev-sorted,sorted-middle-1,rev-sorted-middle-10,random,almost-sorted-.1,sorted,rev-sorted-middle-1,sorted-middle-.1,almost-sorted-10,almost-sorted-1,sorted-middle-10,rev-sorted-end-.1,sorted-end-1,random-dup-10,}
      ,
      xlabel near ticks, x tick label style={rotate=45,anchor=east,font=\tiny},
      ytick distance={200},
      legend style = {
%         legend pos= north east,
        at = {(.87,.97)},
        cells={anchor=west},
        font=\tiny
      },
    ]

      \input{#2}

      \addplot[black, thin, sharp plot,update limits=false]
	coordinates {([normalized]-10,1) ([normalized]100,1)};
% 	node[above] at (axis cs:1950,4.3e7) {Houses};
    \end{axis}
  \end{tikzpicture}
  }

  \plottim{uint64}{d_uint64.tex}

  \clearpage

  \plottim{llstring}{d_llstring.tex}

\end{document}